\begin{mdframed}[backgroundcolor=red!20, frametitlerulewidth=0pt, innertopmargin=-2mm, innerbottommargin=2mm, skipabove=0mm]

\section{Linear Quadratic Regulator LQR}
\end{mdframed}
\begin{minipage}{0.33\textwidth}
    \begin{tcolorbox}[colframe=green!50!black, colback=green!5!white, title=Pros, left=0.5mm, right=0.5mm]
    \begin{itemize}[leftmargin=*]
        \item Optimal control law
        \item Guaranteed stability of the closed-loop system
        \item Online computationally less expensive 
        \item Ideal for linear systems
        \item Flexibility with tuning weights
    \end{itemize}
    \end{tcolorbox}
\end{minipage}
\begin{minipage}{0.33\textwidth}
    \begin{tcolorbox}[colframe=red!50!black, colback=red!5!white, title=Cons, left=0.5mm, right=0.5mm]
    \begin{itemize}[leftmargin=*]
        \item Linear system model 
        \item Model accuracy 
        \item Non-quadratic cost
        \item Constraints 
        \item Robustness is limited
        \item Large memory for finite horizon
    \end{itemize}
    \end{tcolorbox}
\end{minipage}
\begin{minipage}{0.33\textwidth}
    \begin{tcolorbox}[colframe=gray!50!black, colback=gray!5!white, title=Examples, left=0.5mm, right=0.5mm]
    \begin{itemize}[leftmargin=*]
        \item Aircraft autopilot
        \item Path tracking mobile robots
        \item Infinite horizon: persistent tracking in process automation, satellite trajectory
    \end{itemize}
    \end{tcolorbox}
\end{minipage}\\ \\
In general, optimization problems are typically intractable:\\
\begin{minipage}[t]{0.4\textwidth}
 \begin{itemize}
    \item size of $u$ is too large.
    \item require an accurate model of the system.   
 \end{itemize}
\end{minipage}
\begin{minipage}[t]{0.6\textwidth}
 \begin{itemize}
    \item non-convex (between two solutions you don't get another solution).
    \item open-loop nature make them useless in the presence of disturbances. 
 \end{itemize}
\end{minipage}

\subsubsection*{Finite Horizon}
$\Rightarrow$ returns in a sequence of time-varying gain matrices.\\
Tractable optimization problem: Linear dynamics and Quadratic cost.
\begin{itemize}
    \item \textbf{Markovian state} representation:
        $\left\{
        \begin{aligned}
            &\text{Next state is linearly dependent only on current state: }x_{t+1} = A x_t + B u_t. \\
            &\text{Quadratic cost is additive over time: } \sum_t x_t^\top Q x_t + u_t^\top R u_t. \\
            &\text{Initial condition $x_0$ is known.}
        \end{aligned}
        \right.$
    \item The problem is convex and decomposable $\rightarrow$ solvable via \textbf{backward induction} (\textcolor{blue}{\textbf{Bellman Optimality Principle}}: An optimal policy has the property that whatever the initial state and the initial decisions are, the remaining decisions must constitute an optimal policy with regard to the state resulting from the first decision.)
    \begin{align*}
    V_t(x) = \min_{u_t, \ldots, u_{T-1}} &\sum_{s=t}^{T-1} ( \underbrace{x_s^\top Q x_s + u_s^\top R u_s}_{\text{running cost}} ) + \underbrace{x_T^\top S x_T}_{\text{terminal cost}} \\
    \text{s.t.} \quad x_{s+1} &= A x_s + B u_s, \quad s = t, \ldots, T-1 \\
    x_t &= x
    \end{align*}
    where the optimal control sequence is given by 
    \begin{align*}
        u_t &= \underbrace{-(R + B^\top P_{t+1} B)^{-1} B^\top P_{t+1} A}_{\Gamma_t} x_t \\
        P_{t-1} &= Q + A^\top P_t A - A^\top P_t B (R + B^\top P_t B)^{-1} B^\top P_t A, \quad P_T = S
    \end{align*}
\end{itemize}

\textbf{Considerations}
\begin{itemize}
    \item $V_t(x)$ is the value function.
    \item $V_t(x) = x^\top P_t x$ is quadratic, i.e $P \succeq 0$ and symmetric.
    \item $V_0(x_0)$ is the (total) optimal cost.
    \item The result is a linear feedback control law, i.e accounts for disturbances. 
    \item \begin{minipage}[t]{0.45\textwidth}
            \textbf{Offline computation}
            \begin{itemize}
                \item $n \times n$ matrix multiplications ($A$)
                \item $m \times m$ matrix inversion ($R$)
                \item $T$ iterations (horizon)
            \end{itemize}
         \end{minipage}\hfill
         \begin{minipage}[t]{0.45\textwidth}
            \textbf{Online computation}
            \begin{itemize}
                \item Storage of $T$ $m \times n$ matrices ($\Gamma$)
                \item $m \times n$ matrix multiplications ($\Gamma_t x_t$)
            \end{itemize}
        \end{minipage}
\end{itemize}

\newpage

\subsubsection*{Infinite Horizon}
$\Rightarrow$ returns a constant gain matrix.\\
Extension of the LQR to an infinite horizon $T = \infty$ for LTI persistent tracking and regulation problems.
\begin{itemize}
    \item The iteration 
    \begin{equation*}
        P_{t-1} = Q + A^\top P_t A - A^\top P_t B (R + B^\top P_t B)^{-1} B^\top P_t A, \quad P_0 = 0
    \end{equation*}
    converges, as \( t \to \infty \), to a solution \( P_\infty \succ 0 \) of the Algebraic Riccati Equation (ARE):
    \begin{equation*}
        P = Q + A^\top P A - A^\top P B (R + B^\top P B)^{-1} B^\top P A.
    \end{equation*}
    where the optimal feedback control is given by
    \begin{equation*}
        u_t = \underbrace{-(R + B^\top P_{\infty} B)^{-1} B^\top P_{\infty} A}_{\Gamma_\infty} x_t
    \end{equation*}
\end{itemize}

\textbf{Considerations}
\begin{itemize}
    \item The sequence \( P_0, P_{-1}, P_{-2}, \ldots \) is non-decreasing due to Bellman Optimality Principle.
    \item The sequence \( P_0, P_{-1}, P_{-2}, \ldots \) is upper-bounded because it converges towards zero.
    \item $(A,B)$ stabilizable and $(Q^{\frac{1}{2}},A)$ detectable then ARE has a unique $P_{\infty} \succeq 0$ and $A+BK_\infty$ is asy. stable.
    \item In practice: unstable modes need to be weighted in the cost function. 
    \item \begin{minipage}[t]{0.45\textwidth}
            \textbf{Offline computation}
            \begin{itemize}
                \item $n \times n$ matrix multiplications ($A$)
                \item $m \times m$ matrix inversion ($R$)
                \item "$\infty$" iterations (horizon)
            \end{itemize}
         \end{minipage}\hfill
         \begin{minipage}[t]{0.45\textwidth}
            \textbf{Online computation}
            \begin{itemize}
                \item Storage of one $m \times n$ matrix (less memory required!)
                \item $m \times n$ matrix multiplications ($\Gamma_t x_t$)
            \end{itemize}
        \end{minipage}
\end{itemize}
